---
title: "PB_PCA"
output: html_notebook
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2) #I guess this is already in the tidyverse so I don't need to load
```


The polar bear SNPs PCA was run on tassel on the farm. Now those results will be turned into a PCA plot.

Read in Data and Clean
```{r}
pbPCread <- read.table("C:/Users/ebola/ECL243//data_output/PCApbsubfulldata1.txt", comment.char = "", fill=T, sep="\t") #leaving header=true confused it, so I just read it in without, need to get rid of two rows

#pbPCA <- pbPCA[-c(1,2)] this deleted columns, not rows!

pbPCread1 <- pbPCread[-1,]
pbPCA <- pbPCread1[-1,] #this data is almost read

colnames(pbPCA) <- c("taxa", "PC1", "PC2", "PC3")
pbPCA <- pbPCA[-1,] #columns renamed and extra rows removed
```

Plot data
```{r}
ggplot (data = pbPCA) +
  geom_point(mapping = aes(x = PC1, y = PC2)) 
#hooray this gave a PCA plot! Unfortuately, it looks really different from the PCA plot in the paper.

#I want to rearrange the rows to make it easier to plot the east and west greenland bears.



```

I will try adding a column for East or West Greenland bears, then re-plot to see if this shows differences.
```{r}
pbPCA$location <- "West Greenland"
pbPCA$location[1] <- "East Greenland"
pbPCA$location[3] <- "East Greenland"
pbPCA$location[4] <- "East Greenland"
pbPCA$location[5] <- "East Greenland"
pbPCA$location[7] <- "East Greenland"
pbPCA$location[9] <- "East Greenland"
pbPCA$location[11] <- "East Greenland"
```

Now plotting with East and West Greenland separate, ideally with East Greenland having plus shapes and West Greenland having crosses
```{r}

pbPCAplot<- ggplot (data = pbPCA, aes(PC1, PC2)) + #this means we are plotting the PC1 & PC2
  geom_point(aes(shape = location, size = 4)) + #telling it that I want the shapes to change by location
  scale_shape_manual(values = c(3,4)) + #telling it that I want crosses and pluses
  theme(legend.position = "right" ) + #legend
  theme_classic() + #white background
  ggtitle("Polar Bear High-Coverage PCA")  #title
pbPCAplot
#ggsave("pbPCAplot.pdf")

```

Now plot again but, try with PCs as numeric (rather than factor data)

```{r}
pbPCAnumeric <- pbPCA
pbPCAnumeric$PC1 <- as.numeric(as.character(pbPCAnumeric$PC1))
pbPCAnumeric$PC2 <- as.numeric(as.character(pbPCAnumeric$PC2))
str(pbPCAnumeric) #now PC1 is numeric for this data

#plot as numeric but not log
pbPCAplotnumeric<- ggplot (data = pbPCAnumeric, aes(PC1, PC2)) + #this means we are plotting the PC1 & PC2
  geom_point(aes(shape = location)) + #telling it that I want the shapes to change by location
  scale_shape_manual(values = c(3,4)) + #telling it that I want crosses and pluses
  theme(legend.position = "bottom" ) + #legend
  theme_classic() + #white background
  ggtitle("Polar Bear High-Coverage PCA (with numeric data)") #title
pbPCAplotnumeric # this looks good, data is grouped differetly...
ggsave("pbPCAplotnumeric.pdf")
```



