---
title: "polar_bear_data"
output: html_notebook
---

Need to get the data into the correct hapmap format!
First I'll use what Lauren did, then add on some of the stuff below.

Read in the data and unite and fix columns, including adding columns, deleting columns
```{r}
library(tidyr)
library(dplyr)
pbdf1<- read.table("C:/Users/ebola/ECL243/data/polarbearattempt.txt", header=TRUE, comment.char = "", fill=T, sep="\t")
#make the rs column which is a combination of chromo and position
pbdf1 <- unite(pbdf1, rs, c(chromo, position), remove=FALSE)
head (pbdf1)
#make alleles column which is a combination of major and minor
dat3 <- unite (pbdf1, alleles, c(major, minor), remove=FALSE)
head(dat3)
#add in the columns we need, with NAs
dat3$strand <- 0
dat3$assembly <- NA #change this to assembly# in unix, change chromo and position to chrom and pos in unix as well
dat3$center <- NA
dat3$protLSID <- NA
dat3$assayLSID <- NA
dat3$panelLSID <- NA
dat3$QCcode <- NA
head(dat3)
#re-arrange the order of the columns
dat4 <- dat3[c(1,2,3,4,30,31,32,33,34,35,36,5,6,7,8,9,10,11,12,13,14,15,16,17,18,20,22,21,22,23,24,25,26,27,28,29)]
head(dat4)

#cut the columns we don't need
names(dat4) #getting the numbers of the columns we will cut
dat5 <- dat4[,-12] 
names(dat5)
dat6 <- dat5[,-4]
names(dat6)
dat7 <- dat6[,-12]
dat8 <- dat7[,-12]
dat9 <- dat8[,-12]

head (dat9)
names(dat9)
dat10 <- dat9[,-12]
names(dat10)
dat11 <- dat10[,-12]
names(dat11)
head(dat11)
pbdf3 <- dat11
head(pbdf3)

```

Ok, just need to fix the order, change _ to /, then fix strand and we are good.

we need the order to be: rs#    alleles    chrom     pos     strand    assembly#    center    protLSID    assayLSID    panelLSID    QCcode    sample1 etc.

```{r}
names(pbdf3)
pbdf4 <-pbdf3[c(1,11,2,3,4,5,6,7,8,9,10,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29)]
head(pbdf4)

#use gsub to replace one symbol with another in the alleles column


pbdf4$alleles <- gsub("_", "/", pbdf4$alleles)
head(pbdf4)

#add in the pluses for strand
pbdf4$strand <- rep('+', nrow(pbdf4))
head(pbdf4)

#double check "scaffold" is gone
pbdf4$chromo <- gsub("scaffold", "", pbdf4$chromo)

```


Ok, the data is clean in R! Now we export, then Lauren will fix some things.

```{r}

#write.csv(pbdf4, file = "polarberSNPclean.csv", row.names = F) didn't see to write a csv...

#write.table(pbdf4, "C:/Users/ebola/ECL243/data/polarbearSNPclean.txt", quote = F, sep="\t", row.names = F)



```


